---
title: "Feux agricoles et pastoraux"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Initialization

1.  Setup layout parameters

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.  Setup global options

```{r, include=FALSE}
# options(java.parameters = "- Xmx1024m") #Allow to increase memory limit
```

3.  Load required packages

```{r, message=FALSE}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(readxl)
library(lubridate)
library(annotater) # Annotate Package Load Calls, [github::luisDVA/annotater] v0.1.3
library(sf) # Simple Features for R
library(tmap) # Thematic Maps
library(leaflet) # Create Interactive Web Maps with the JavaScript 'Leaflet'
library(tmaptools) # Thematic Map Tools
library(snakecase) # Convert Strings into any Case
library(ggridges)
library(pilot)
library(ggdist)
library(inspectdf)
library(ggforce)


```

4.  Get data

```{r}
agri_fire <- read_excel("data/Base feux agricoles 2001-2021-31122021.xlsx", sheet = 1)
pasto_fire <- read_excel("data/Base feux agricoles 2001-2021-31122021.xlsx", sheet = 2)
commune_gps <- read_csv2("data/commune_gps_tidied.csv")
```

# Function created

```{r}
bubble.map <- function(group_type = "agri_fire") {
  tmap_mode("view")
  fire_sf %>%
    filter(groups == group_type) %>%
    tm_shape() +
    tm_dots(
      size = "surface_ha", col = "surface_ha", palette = c("#ffffcc", "#ffeda0", "#fed976", "#feb24c", "#fd8d3c", "#fc4e2a", "#e31a1c", "#b10026"),
      title = "Burned area (ha)", alpha = 0.8, n = 8, scale = 2
    ) -> bubble_map
  print(bubble_map)
  return(bubble_map)
}

save.map <- function(map_type = "bubble", group_type = "agri_fire", directory = "local") {
  title <- "BA_"
  if (map_type == "bubble") {
    suffix <- ".html"
    title <- paste0(title, "bubble_")
    if (group_type == "agri_fire") {
      title <- paste0(title, "agri")
    } else {
      title <- paste0(title, "pasto")
    }
  } else {
    suffix <- ".png"
    tmap_mode("plot")
    title <- paste0(title, "Facet")
  }
  if (directory == "local") {
    filename <- paste0("output_data/", title)
  } else {
    filename <- paste0("M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/", title)
  }
  filename <- paste0(filename, suffix)

  bubble_map <- eval(parse(text = title))
  tmap_save(bubble_map, filename)
}
```

# Data Cleaning and Tidying

1.  Data Cleaning

-   Agricultural fire data

```{r}
agri_fire <- as_tibble(agri_fire)
agri_fire %>%
  filter(!mois == "") %>%
  select(année:rmq) -> agri_fire
```

-   Pastoral fire data

```{r}
pasto_fire <- as_tibble(pasto_fire)
pasto_fire %>%
  filter(!mois == "") %>%
  select(année:rmq) -> pasto_fire
```

2.  Data Tidying

-   Agricultural fire data

```{r}
# Rename columns
clean_names(agri_fire) -> agri_fire
agri_fire %>%
  rename(code_commune = `x4`) -> agri_fire

# Filter miswritten rows or change its value
agri_fire %>%
  filter(is.na(surface_ha)|!surface_ha==".0°4"&!surface_ha=="1,5")%>%
  mutate(surface_ha = na_if(surface_ha, "na"),
         masse_tonne_matiere_seche = na_if(masse_tonne_matiere_seche, "na"),
         type_vegetation = na_if(type_vegetation, "na"),
         lieu_de_stockage = recode(lieu_de_stockage, "na" = "not_mentioned"),
         lieu_de_stockage = na_if(lieu_de_stockage, "NA"))-> agri_fire

# Set class of each column
agri_fire %>%
  mutate(
    annee = as.integer(annee),
    mois = as.integer(mois),
    jour = as.integer(jour),
    code_commune = as.character(code_commune),
    commune = as.character(commune),
    surface_ha = as.numeric(surface_ha),
    type_vegetation = as.character(type_vegetation),
    lieu_de_stockage=as.character(lieu_de_stockage),
    masse_tonne_matiere_seche = as.numeric(masse_tonne_matiere_seche),
    source = as.factor(source),
    rmq =as.character(rmq)
  ) -> agri_fire
agri_fire %>%
  mutate(
    type_vegetation = iconv(type_vegetation, from = "UTF-8", to = "ASCII//TRANSLIT"),
    type_vegetation = to_snake_case(type_vegetation),
    type_vegetation = as.factor(type_vegetation)
  ) -> agri_fire

# Create new columns
agri_fire %>%
  mutate(
    date = make_date(annee, mois, jour),
    datemj = make_date(month = mois, day = jour)
  ) -> agri_fire
agri_fire %>%
  mutate(datemj = str_sub(datemj, -5)) -> agri_fire

# agri_fire%>%
#   filter(surface_ha=="na")%>%
#   count->na1
# test%>%
#   filter(is.na(surface_ha))%>%
#   count->na2
# na1==na2
#
# seqna1<-which(agri_fire$surface_ha=="na")
# seqna2<-which(is.na(test$surface_ha))
# tabcomp<-cbind(seqna1,seqna2)
# tabcomp<-as_tibble(tabcomp)
# tabcomp%>%
#   mutate(diff=(seqna1-seqna2))->tabcomp
# tabcomp%>%
#   filter(diff!=0)

# 370,1397,3225
```

-   Pastoral fire data

```{r}
# Rename columns
clean_names(pasto_fire) -> pasto_fire

# Filter miswritten rows or change its value
pasto_fire %>%
  filter(!surface_ha=="> 100" & !surface_ha=="> 101"&!jour=="?"& !jour == "1'")%>%
  mutate(surface_ha = na_if(surface_ha, "NA"),
         surface_ha = na_if(surface_ha, "na"),
         type_vegetation = na_if(type_vegetation, "na")) -> pasto_fire

# Set class of each column
pasto_fire %>%
  mutate(
    annee = as.integer(annee),
    mois = as.integer(mois),
    jour = as.integer(jour),
    code_commune = as.character(code_commune),
    commune = as.character(commune),
    surface_ha = as.numeric(surface_ha),
    type_vegetation = as.factor(type_vegetation),
    source = as.factor(source)
  ) -> pasto_fire
pasto_fire %>%
  mutate(
    type_vegetation = iconv(type_vegetation, from = "UTF-8", to = "ASCII//TRANSLIT"),
    type_vegetation = to_snake_case(type_vegetation),
    type_vegetation = as.factor(type_vegetation)
  ) -> pasto_fire

# Create new columns
pasto_fire %>%
  mutate(
    date = make_date(annee, mois, jour),
    datemj = make_date(month = mois, day = jour)
  ) -> pasto_fire
pasto_fire %>%
  mutate(datemj = str_sub(datemj, -5)) -> pasto_fire

# pasto_fire%>%
#   filter(surface_ha=="na")%>%
#   count->na1
# test%>%
#   filter(is.na(surface_ha))%>%
#   count->na2
# na1==na2
#
# seqna1<-which(pasto_fire$surface_ha=="na")
# seqna2<-which(is.na(test$surface_ha))
# tabcomp<-cbind(seqna1,seqna2)
# tabcomp<-as_tibble(tabcomp)
# tabcomp%>%
#   mutate(diff=(seqna1-seqna2))->tabcomp
# tabcomp%>%
#   filter(diff!=0)

# 370,1397,3225
```

-   Commune GPS data

```{r}
# Filter miswritten rows or change its value
commune_gps%>%
  mutate(code_insee=if_else(str_length(code_insee)==4, paste0(0,code_insee),code_insee))-> commune_gps

# Set class of each column
commune_gps %>%
  mutate(code_insee = as.character(code_insee)) -> commune_gps

# Create new columns
commune_gps %>%
  separate(geo_point_2d, c("lat", "lon"), sep = ",") %>%
  mutate(
    lon = as.numeric(lon),
    lat = as.numeric(lat)
  ) -> commune_gps
```

3.  Data merging

```{r}
fire_mix <- bind_rows("agri_fire" = agri_fire, "pasto_fire" = pasto_fire, .id = "groups")
fire <- left_join(fire_mix, commune_gps, by = c("code_commune" = "code_insee"))
fire %>%
  filter(!is.na(lat)) -> fire
# fire%>%
# filter(is.na(lat))%>%
#   count(code_commune)->unavailable_loc
```

4.  Creating other format files

```{r}
fire_sf <- st_as_sf(fire,
  coords = c("lon", "lat"),
  crs = "+proj=longlat +datum=WGS84 +nodefs"
)
```

# Data exploration

1.  Graphic exploration

```{r}
fire%>%
  inspect_na()%>%
  show_plot()+
  theme_pilot()->NA_Variable_barplot
fire%>%
  inspect_num()%>%
  show_plot()+
  theme_pilot()->Variable_histogram
fire%>%
  inspect_cat()%>%
  show_plot()+
  theme_pilot()+
  theme(legend.position="none")->Variable_Frequency_barplot
fire %>%
  ggplot() +
  aes(x = date, color = groups, fill = groups) +
  geom_density(alpha = 0.5, adjust = 1 / 10) +
  labs(color = "Type of fire event",
       fill = "Type of fire event",
       x = "Year",
       y = "Number of fire events"
      ) +
  theme_pilot() +
  scale_color_pilot(labels = c("Agricultural fire", "Pastoral fire"))+
  scale_fill_pilot(labels = c("Agricultural fire", "Pastoral fire"))-> NbFire_Date_density
fire %>%
  group_by(groups, annee, mois) %>%
  count() %>%
  group_by(groups, mois) %>%
  summarise(mean = mean(n)) %>%
  ggplot() +
  aes(x = as.factor(mois), fill = groups, color = groups, y = mean) +
  geom_bar(stat = "identity") +
  ggtitle("Mean Number of fire events per month from 2001 to 2021") +
  labs(color = "Type of fire event", 
       fill = "Type of fire event",
       x = "Month",
       y = "Number of fire events") +
    theme_pilot() +
  scale_color_pilot(labels = c("Agricultural fire", "Pastoral fire"))+
  scale_fill_pilot(labels = c("Agricultural fire", "Pastoral fire"))-> MeanNbFire_Month_barplot 
fire %>%
  mutate(type_vegetation_lump = fct_lump(type_vegetation, prop = .004, )) %>%
  ggplot() +
  aes(x = fct_infreq(type_vegetation_lump), color = groups, fill = groups) +
  geom_bar() +
  coord_flip() +
  xlab("Vegetation type") +
  ylab("Count") +
  ggtitle("Frequence of each vegetation type") +
  labs(color = "Type of fire event", fill = "Type of fire event") +
  theme_pilot() +
  scale_color_pilot(labels = c("Agricultural fire", "Pastoral fire"))+
  scale_fill_pilot(labels = c("Agricultural fire", "Pastoral fire")) -> VegetType_Nb_barplot
fire %>%
  mutate(code_commune_lump = fct_lump(code_commune, prop = .0009)) %>%
  filter(code_commune_lump != "Other") %>%
  ggplot() +
  aes(x = fct_infreq(commune.y), color = groups, fill = groups) +
  geom_bar() +
  coord_flip() +
  xlab("Municipality") +
  ylab("Count") +
  ggtitle("Most affected municipalities over the study period, (corresponding to 0.09% of data)") +
  labs(color = "Type of fire event", fill = "Type of fire event",
       x="Number of fire events") +
  theme_pilot() +
  scale_color_pilot(labels = c("Agricultural fire", "Pastoral fire"))+
  scale_fill_pilot(labels = c("Agricultural fire", "Pastoral fire")) -> City_Nb_barplot

level_of_fire <- as.tibble(levels(as.factor(fire$type_vegetation)))
# level_of_fire%>%
# filter(str_detect(value, "chaume") & str_detect(value, "ble"))
word_in_fire <- str_extract_all(level_of_fire, "[a-z]+")
as.tibble(word_in_fire[[1]]) %>%
  mutate(value_lump = fct_lump(value, prop = .003, )) %>%
  ggplot() +
  aes(x = fct_infreq(value_lump)) +
  geom_bar() +
  coord_flip() +
  xlab("Word inside vegetation type") +
  ylab("Count") +
  ggtitle("Frequence of each word among the 402 vegetation type, nb words = 818") +
  theme_pilot() +
  scale_color_pilot()+
  scale_fill_pilot() -> Words_Nb_barplot
fire %>%
  mutate(type_vegetation_lump = fct_lump(type_vegetation, prop = .004, )) %>%
  filter(type_vegetation_lump != "Other" & surface_ha < 50) %>%
  ggplot() +
  aes(x = type_vegetation_lump, y = surface_ha) +
  stat_halfeye() +
  coord_flip() +
  xlab("Vegetation type") +
  ylab("Burned Area (ha)") +
  ggtitle("Burned Area of most common vegetation type, include only fires<50 ha") +
  theme_pilot() +
  scale_color_pilot()+
  scale_fill_pilot() -> VegetType_Surface_distplot
fire %>%
  ggplot() +
  aes(x = datemj, y = groups, group = groups, color = groups, fill = groups) +
  geom_density_ridges2() +
  xlab("Date") +
  ylab("Number of fire events") +
  labs(color = "Type of fire event", fill = "Type of fire event") +
  theme_pilot() +
  scale_color_pilot(labels = c("Agricultural fire", "Pastoral fire"))+
  scale_fill_pilot(labels = c("Agricultural fire", "Pastoral fire")) +
  facet_wrap(~annee, ncol = 1, strip.position = "right") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.spacing = unit(0, "lines"),
    axis.text.x=element_blank()
  ) -> NbFire.Year_JJMM_density
fire%>%
  mutate(groups=as.numeric(as.factor(groups)))%>%
  ggplot()+
  aes(x =groups,
      y=surface_ha,
      color = groups) +
    geom_jitter(alpha = 0.5,
                width=.2) +
    labs(title = "Distribution of Burned area ~ type of fire",
         subtitle = "Dark blue = Agricultural fire, Light blue = Pastoral fire",
         x = "Fire type",
         y = "Burned area (ha)") +
  theme(legend.position = "none",
        axis.text.x = element_blank())+
  facet_zoom(ylim = c(0, 200), xlim= c(0.8,2.2),zoom.size = 0.5)-> BA_Group_jitterplot
```

2.  Spatial exploration

```{r}
tmap_mode("view")
bubble.map("agri_fire") -> BA_bubble_agri
bubble.map("pasto_fire") -> BA_bubble_pasto

# fire%>%
#   ggplot()+
#   aes(x=lon,y=lat)+
#   geom_density_2d_filled()+
#   facet_wrap(~groups)+
#     xlab("Vegetation type")+
#   ylab("Burned Area (ha)")+
#   ggtitle("Burned Area of most common vegetation type, include only fires<50 ha")+
#   theme_bw()
# fire_sf_density<-density(x = fire_sf$code_commune)
# tm_shape(fire_sf)+

tmap_mode("plot")
osm_fire_sf <- read_osm(fire_sf, ext = 1.1)
tm_shape(osm_fire_sf) +
  tm_rgb() +
  tm_shape(fire_sf) +
  tm_tiles("OpenStreetMap") +
  tm_dots(
    col = "groups",
    title = "Burned area (ha)",
    palette = pilot_color()[c(1,length(pilot_color()))],
    alpha = 0.8,
    n = 8,
    scale = 2,
    legend.show = F
  ) +
  tm_facets(by = "groups", free.coords = F) +
  tm_layout(panel.labels = c("Agricultural fire events", "Pastoral fire events")) -> BA_Facet
print(BA_Facet)

fire_sf%>%
  ggplot()+
  aes(color=groups,size=surface_ha)+
  geom_sf(alpha=0.6)+
  theme_pilot() +
  scale_color_pilot()+
  scale_size_area(max_size = 20)+
    labs(color = "Type of fire event",
       size = "Burned area (ha)"
      )+
  borders("france")-> BA_Group_bubblemap
mapview(fire_sf, zcol="groups")
mapview(fire_sf, cex="surface_ha",)
```

# Export data

1.  Export data

```{r}
write_csv2(fire,"Agricultural-Pastoral_Fire.csv")
```

2.  Export figures

```{r}
ggsave(NA_Variable_barplot, filename = "NA_Variable_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(NA_Variable_barplot, filename = "NA_Variable_barplot.png", path = "output_data")
ggsave(Variable_histogram, filename = "Variable_histogram.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(Variable_histogram, filename = "Variable_histogram.png", path = "output_data")
ggsave(Variable_Frequency_barplot, filename = "Variable_Frequency_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(Variable_Frequency_barplot, filename = "Variable_Frequency_barplot.png", path = "output_data")
ggsave(NbFire_Date_density, filename = "NbFire_Date_density.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(NbFire_Date_density, filename = "NbFire_Date_density.png", path = "output_data")
ggsave(MeanNbFire_Month_barplot, filename = "MeanNbFire_Month_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(MeanNbFire_Month_barplot, filename = "MeanNbFire_Month_barplot.png", path = "output_data")
ggsave(VegetType_Nb_barplot, filename = "VegetType_Nb_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(VegetType_Nb_barplot, filename = "VegetType_Nb_barplot.png", path = "output_data")
ggsave(City_Nb_barplot, filename = "City_Nb_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(City_Nb_barplot, filename = "City_Nb_barplot.png", path = "output_data")
ggsave(Words_Nb_barplot, filename = "Words_Nb_barplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(Words_Nb_barplot, filename = "Words_Nb_barplot.png", path = "output_data")
ggsave(VegetType_Surface_distplot, filename = "VegetType_Surface_distplot.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(VegetType_Surface_distplot, filename = "VegetType_Surface_distplot.png", path = "output_data")
ggsave(NbFire.Year_JJMM_density, filename = "NbFire.Year_JJMM_density.png", path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(NbFire.Year_JJMM_density, filename = "NbFire.Year_JJMM_density.png", path = "output_data")
tmap_mode("view")
for (group_type in c("agri_fire", "pasto_fire")) {
  for (directory in c("local", "mouillot")) {
    save.map(map_type = "bubble", group_type = group_type, directory = directory)
  }
}
tmap_mode("plot")
for (directory in c("local", "mouillot")) {
  save.map(map_type = "BA", directory = directory)
}
```
