---
title: "Feux agricoles et pastoraux"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Initialization

1.  Setup layout parameters

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.  Setup global options

```{r, include=FALSE}
#options(java.parameters = "- Xmx1024m") #Allow to increase memory limit
```

3.  Load required packages

```{r, message=FALSE}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(readxl) # Read Excel Files, CRAN v1.3.1 
library(data.table) # Extension of `data.frame`
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(lubridate) # Make Dealing with Dates a Little Easier
library(annotater) # Annotate Package Load Calls, [github::luisDVA/annotater] v0.1.3
library(sf)
library(tmap)
library(leaflet)
library(tmaptools)
library(snakecase)

```

4.  Get data

```{r}
agri_fire<-read_excel("data/Base-feux-agricoles-pastoraux2001-2021-10092021.xlsx",sheet = 1)
pasto_fire<-read_excel("data/Base-feux-agricoles-pastoraux2001-2021-10092021.xlsx",sheet = 2)
commune_gps<-read_csv2("data/commune_gps_tidied.csv")
```

# Function created

```{r}
bubble.map<-function(group_type="agri_fire"){
  tmap_mode("view")
fire_sf%>%
  filter(groups==group_type)%>%
tm_shape()+
  tm_dots(size="surface_ha",col="surface_ha",palette=c('#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#b10026'),
          title="Burned area (ha)",alpha=0.8,n=8,scale=2)->bubble_map
print(bubble_map)
return(bubble_map)
}

save.map<-function(map_type="bubble",group_type="agri_fire",directory="local"){
  
  title<-"BA_"
  if(map_type=="bubble"){
    suffix<-".html"
    title<-paste0(title,"bubble_")
    if(group_type=="agri_fire"){
      title<-paste0(title,"agri")}
    else{
      title<-paste0(title,"pasto")
    }
  }
  else{
    suffix<-".png"
    tmap_mode("plot")
    title<-paste0(title,"Facet")
  }
if(directory=="local"){
  filename<-paste0("output_data/",title)
}
else{
  filename<-paste0("M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/",title)
}
filename<-paste0(filename,suffix)

bubble_map<-eval(parse(text = title))
tmap_save(bubble_map,filename)
}
```

# Data Cleaning and Tidying

1.  Data Cleaning

-   Agricultural fire data

```{r}
agri_fire<-as_tibble(agri_fire)
agri_fire%>%
  filter(!mois=="")%>%
  select(!rmq&!`...10`)->agri_fire

```

-   Pastoral fire data

```{r}
pasto_fire<-as_tibble(pasto_fire)
pasto_fire%>%
  filter(!mois=="")%>%
  select(!rmq&!`...10`)->pasto_fire
```

2.  Data Tidying

-   Agricultural fire data

```{r}
clean_names(agri_fire)->agri_fire
agri_fire%>%
  rename(code_commune=`x4`)->agri_fire

agri_fire%>%
  slice(c(1:369,371:1396,1398:nrow(agri_fire)))%>%
  na.omit(code_commune)-> agri_fire

agri_fire%>%
  mutate(annee=as.integer(annee),
         mois=as.integer(mois),
         jour=as.integer(jour),
         code_commune=as.character(code_commune),
         commune=as.character(commune),
         surface_ha=as.numeric(surface_ha),
         type_vegetation=as.character(type_vegetation),
         source=as.factor(source))->agri_fire

# agri_fire%>%
#   filter(surface_ha=="na")%>%
#   count->na1
# test%>%
#   filter(is.na(surface_ha))%>%
#   count->na2
# na1==na2
# 
# seqna1<-which(agri_fire$surface_ha=="na")
# seqna2<-which(is.na(test$surface_ha))
# tabcomp<-cbind(seqna1,seqna2)
# tabcomp<-as_tibble(tabcomp)
# tabcomp%>%
#   mutate(diff=(seqna1-seqna2))->tabcomp
# tabcomp%>%
#   filter(diff!=0)

#370,1397,3225

agri_fire%>% 
  mutate(date = make_datetime(annee, mois, jour))->agri_fire
agri_fire%>%
  mutate(type_vegetation=replace(type_vegetation, type_vegetation=="na",NA),
                                 type_vegetation=iconv(type_vegetation,from="UTF-8",to="ASCII//TRANSLIT"),
                                 type_vegetation=to_snake_case(type_vegetation),
         type_vegetation=as.factor(type_vegetation))->agri_fire
```

-   Pastoral fire data

```{r}
clean_names(pasto_fire)->pasto_fire
pasto_fire%>%
  select(annee:source)->pasto_fire

pasto_fire%>% 
  slice(c(1:61,64:76,78:817,819:821,823:945,947:nrow(pasto_fire)))-> pasto_fire

pasto_fire%>%
  mutate(annee=as.integer(annee),
         mois=as.integer(mois),
         jour=as.integer(jour),
         code_commune=as.character(code_commune),
         commune=as.character(commune),
         surface_ha=as.numeric(surface_ha),
         type_vegetation=as.factor(type_vegetation),
         source=as.factor(source))->pasto_fire

# pasto_fire%>%
#   filter(surface_ha=="na")%>%
#   count->na1
# test%>%
#   filter(is.na(surface_ha))%>%
#   count->na2
# na1==na2
# 
# seqna1<-which(pasto_fire$surface_ha=="na")
# seqna2<-which(is.na(test$surface_ha))
# tabcomp<-cbind(seqna1,seqna2)
# tabcomp<-as_tibble(tabcomp)
# tabcomp%>%
#   mutate(diff=(seqna1-seqna2))->tabcomp
# tabcomp%>%
#   filter(diff!=0)

#370,1397,3225

pasto_fire%>% 
  mutate(date = make_datetime(annee, mois, jour))->pasto_fire
pasto_fire%>%
  mutate(type_vegetation=replace(type_vegetation, type_vegetation=="na",NA),
                                 type_vegetation=iconv(type_vegetation,from="UTF-8",to="ASCII//TRANSLIT"),
                                 type_vegetation=to_snake_case(type_vegetation),
         type_vegetation=as.factor(type_vegetation))->pasto_fire
```

-   Commune GPS data

```{r}
commune_gps%>%
  mutate(code_insee=as.character(code_insee))->commune_gps
commune_gps%>%
  separate(geo_point_2d,c("lat","lon"),sep=",")%>%
  mutate(lon=as.numeric(lon),
         lat=as.numeric(lat))->commune_gps
```

3.  Data merging

```{r}
fire_mix<-bind_rows("agri_fire"=agri_fire,"pasto_fire"=pasto_fire,.id="groups")
fire<-left_join(fire_mix,commune_gps,by=c("code_commune"="code_insee"))
fire%>%
  filter(!is.na(lat))->fire
```

4.  Creating other format files

```{r}
fire_sf<-st_as_sf(fire,
               coords = c("lon","lat"),
               crs="+proj=longlat +datum=WGS84 +nodefs")
```

# Data exploration

1.  Graphic exploration

```{r}
fire%>%
  ggplot()+
  aes(x=date,color=groups,fill=groups)+
  geom_density(alpha=0.2,adjust=1/10)+
  xlab("Date")+
  ylab("Number of fire events")+
  labs(color="Type of fire event",fill="Type of fire event")+
  scale_color_hue(labels=c("Agricultural fire","Pastoral fire"))+
  scale_fill_hue(labels=c("Agricultural fire","Pastoral fire"))+
  theme_bw()->NbFire_Date_density
fire%>%
  group_by(groups,annee,mois)%>%
  count()%>%
  group_by(groups,mois)%>%
  summarise(mean=mean(n))%>%
  ggplot()+
  aes(x=as.factor(mois),fill=groups,color=groups,y=mean)+
  geom_bar(stat="identity",alpha=0.2)+
  xlab("Month")+
  ylab("Number of fire events")+
  ggtitle("Mean Number of fire events per month from 2001 to 2021")+
  labs(color="Type of fire event",fill="Type of fire event")+
  scale_color_hue(labels=c("Agricultural fire","Pastoral fire"))+
  scale_fill_hue(labels=c("Agricultural fire","Pastoral fire"))+
  theme_bw()->MeanNbFire_Month_barplot
fire%>%
  mutate(type_vegetation_lump=fct_lump(type_vegetation,prop=.004,))%>%
  ggplot()+
  aes(x=fct_infreq(type_vegetation_lump),color=groups,fill=groups)+
  geom_bar()+
  coord_flip()+
  xlab("Vegetation type")+
  ylab("Count")+
  ggtitle("Frequence of each vegetation type")+
  labs(color="Type of fire event",fill="Type of fire event")+
  scale_color_hue(labels=c("Agricultural fire","Pastoral fire"))+
  scale_fill_hue(labels=c("Agricultural fire","Pastoral fire"))+
  theme_bw()->VegetType_Nb_barplot
fire%>%
  mutate(code_commune_lump=fct_lump(code_commune,prop=.001))%>%
  filter(code_commune_lump!="Other")%>%
  ggplot()+
  aes(x=fct_infreq(commune.y),color=groups,fill=groups)+
  geom_bar()+
  coord_flip()+
  xlab("City")+
  ylab("Count")+
  ggtitle("Most affected cities, corresponding to 0.1% of data")+
  labs(color="Type of fire event",fill="Type of fire event")+
  scale_color_hue(labels=c("Agricultural fire","Pastoral fire"))+
  scale_fill_hue(labels=c("Agricultural fire","Pastoral fire"))+
  theme_bw()->City_Nb_barplot
fire%>%
  filter(str_count(type_vegetation))
level_of_fire<-as.tibble(levels(fire$type_vegetation))
level_of_fire%>%
    filter(str_detect(value,'chaume')&str_detect(value,'ble'))
str_count(level_of_fire$value)
str_split(level_of_fire,"_")
word_in_fire<-str_extract_all(level_of_fire, "[a-z]+")
as.tibble(word_in_fire[[1]])%>%
  mutate(value_lump=fct_lump(value,prop=.003,))%>%
  ggplot()+
  aes(x=fct_infreq(value_lump))+
  geom_bar()+
  coord_flip()+
  xlab("Word inside vegetation type")+
  ylab("Count")+
  ggtitle("Frequence of each word among the 402 vegetation type, nb words = 818")+
  theme_bw()->Words_Nb_barplot
```

2.  Spatial exploration

```{r}
tmap_mode("view")
bubble.map("agri_fire")->BA_bubble_agri
bubble.map("pasto_fire")->BA_bubble_pasto

tmap_mode("plot")
osm_fire_sf<-read_osm(fire_sf,ext=1.1)
tm_shape(osm_fire_sf)+
  tm_rgb()+
tm_shape(fire_sf)+
tm_tiles("OpenStreetMap")+
    tm_dots(col="groups",
          title="Burned area (ha)",
          palette=c("#2ca25f","#2b8cbe"),
          alpha=0.8,
          n=8,
          scale=2,
          legend.show =F 
          )+
  tm_facets(by="groups",free.coords = F)+
	tm_layout(panel.labels = c("Agricultural fire events", "Pastoral fire events"))->BA_Facet
print(BA_Facet)
```

# Export data

```{r}
ggsave(NbFire_Date_density,filename = "NbFire_Date_density.png",path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(NbFire_Date_density,filename = "NbFire_Date_density.png",path = "output_data")
ggsave(MeanNbFire_Month_barplot,filename = "MeanNbFire_Month_barplot.png",path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(MeanNbFire_Month_barplot,filename = "MeanNbFire_Month_barplot.png",path = "output_data")
ggsave(VegetType_Nb_barplot,filename = "VegetType_Nb_barplot.png",path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(VegetType_Nb_barplot,filename = "VegetType_Nb_barplot.png",path = "output_data")
ggsave(City_Nb_barplot,filename = "City_Nb_barplot.png",path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(City_Nb_barplot,filename = "City_Nb_barplot.png",path = "output_data")
ggsave(Words_Nb_barplot,filename = "Words_Nb_barplot.png",path = "M:/Departements/EF/EqFORECAST/mouillot/Feux-agricoles-pastoraux/output-data/")
ggsave(Words_Nb_barplot,filename = "Words_Nb_barplot.png",path = "output_data")
tmap_mode("view")
for(group_type in c("agri_fire","pasto_fire")){
  for(directory in c("local","mouillot")){
    save.map(map_type = "bubble",group_type = group_type,directory = directory)
    }
}
tmap_mode("plot")
for(directory in c("local","mouillot")){
save.map(map_type = "BA",directory=directory)
}
```
